---
title: "Homework 1"
author: "Andrew Shao (NetID: as13381)"
output: pdf_document
bibliography: references.bib
nocite:
  - '@*'
header-includes: 
  - \usepackage{tikz}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Question 1 (0.5 pt): Generate the following code chunk.
```{r}
suppressMessages(library(tidyverse))
library(dslabs)
data(us_contagious_diseases)
```

## Question 2 (1 pt): Use the following R code to produce the following bar chart for the dataset `us_contagious_diseases`. Choose an appropriate chunk option to add the caption “`The number of observations for each disease`” to the figure.
```{r, fig.cap='The number of observations for each disease'}
ggplot(data = us_contagious_diseases) + geom_bar(mapping = aes(x = disease))
```

## Question 3 (1 pt): The R code `head(us_contagious_diseases)` outputs the first 6 rows of the dataset `us_contagious_diseases`. Use `knitr::kable` to make the following table for the output of `head(us_contagious_diseases)`, with the caption “`The first 6 rows of us_contagious_diseases`”. Choose an appropriate chunk option to hide the code chunk.
```{r, tab.cap='The first 6 rows of us_contagious_diseases', echo=FALSE}
knitr::kable(head(us_contagious_diseases), "pipe")
```

## Question 4 (0.5 pt): Write the following sentence and formula.
Compute the rate by the following formula:
\[rate = cout/poluation.\]

## Question 5 (0.5 pt): Write the following sentence in which the numbers are produced by inline R code.
The `us_contagious_diseases` data has `r nrow(us_contagious_diseases)` rows and `r ncol(us_contagious_diseases)` columns.

## Question 6 (0.5 pt): Write the following sentence and include a hyperlink to the website address.
The `us_contagious_diseases` data is available in the R package `dslabs` (https://rdrr.io/cran/dslabs/man/us_contagious_diseases.html).

## Question 7 (1 pt): Write the following sentence. Cite the article and generate the reference list by using BibTex.
Please cite the article Van Panhuis et al. (2013) for the dataset `us_contagious_diseases`.

# References