---
title: "Homework 3"
author: "Andrew Shao"
date: "2024-09-25"
output: pdf_document
header-includes: 
  - \usepackage{tikz}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.16.3 Q1
```{r}
sum(!is.na(as.Date(paste0(1900:2021, '-02-29'), format = '%Y-%m-%d')))
```

## 2.16.3 Q2
-10 years in days; R assumes 2 digit years to be in the 2000's if between 00 and 68 otherwise 1900's so `x` is really January 1, 2069 while `y` is December 31, 1968.
```{r}
x <- as.Date("69-01-01", format = "%y-%m-%d")
y <- as.Date("68-12-31", format = "%y-%m-%d")
x - y
```

## 2.16.3 Q3
```{r}
q3 <- as.Date('21-02-14', format = '%y-%m-%d')
q3 + 1000
```

## 2.16.3 Q4
```{r}
q4 <- as.POSIXct('21-07-04 08:15:00', format = '%y-%m-%d %H:%M:%S')
q4 +
  as.difftime(365, units = 'days') +
  as.difftime(2, units = 'days') +
  as.difftime(3, units = 'hours') +
  as.difftime(4, units = 'mins') +
  as.difftime(5, units = 'secs')
```

## 3.2.5 Q1
```{r}
a <- array(1:24, c(2, 3, 4))
apply(a, 2, mean)
```

## 3.2.5 Q2
```{r}
apply(a, c(1, 3), quantile, 0.25)
```

## 3.2.5 Q3
```{r}
apply(a, c(1, 3), quantile, c(0.25, 0.75))
```

## 3.4.4 Q1
```{r}
library(tidyverse)
animal <- rep(c("sheep", "pig"), c(3,3))
weight <- c(110, NA, 140, NA, 300, 800)
condition <- c("excellent", "good", NA, "excellent", "good", "average")
healthy <- c(rep(TRUE, 5), FALSE)
my_tibble <- tibble(animal, weight, condition, healthy)
my_data_frame <- data.frame(animal, weight, condition, healthy)
my_tibble
```
```{r}
add_row(my_tibble, animal = "pig", weight = 900, condition = 'average', healthy = FALSE)
```

## 3.4.4 Q2
`my_tibble[, 1]` should return a `tibble` of size 6 $\times$ 1 while `my_data_frame[, 1]` should return a vector. To reproduce `my_data_frame[, 1]` you should add `drop = TRUE` to `my_tibble[, 1]`.

## 3.5.7 Q1
```{r}
dig_num <- 1:6
ani_char <- c("sheep", "pig", "monkey", "pig", "monkey")
x_mat <- matrix(1:12, nrow = 3, ncol = 4)
my_list <- list(num = dig_num, char = ani_char, mat = x_mat)
my_list
```
`my_list[2]` returns the sublist of element 2 of `my_list` while `my_list[3]` returns the sublist of element 3.

## 3.5.7 Q2
`my_list[2:3]` returns the sublist containing elements 2 and 3 of `my_list` while `my_list[[2:3]]` returns the third element of the second element of `my_list`.

## 3.5.7 Q3
```{r}
sapply(my_list, length)
```

## 3.5.7 Q4
```{r}
my_list$mat[3, ] * 10
```
```{r}
my_list
```

## 3.6.5 Q1
```{r}
x <- c(NA, NULL, Inf, NaN)
```
3
```{r}
length(x)
```

## 3.6.5 Q2
class is `numeric` and storage type is `double`
```{r}
class(x)
```
```{r}
typeof(x)
```

## 3.6.5 Q3
`x + 1` returns `c(NA, Inf, NaN)`. \
Adding 1 to the first element returns `NA` because adding to a `NA` returns `NA` because the values is not available. \
Adding 1 to the second element effectively returns nothing because `NULL` has length 0 so adding to a length 0 object returns another object of length 0. \
Adding 1 to the third element returns `Inf` because `Inf` represents infinity and mathematically, adding 1 to infinity is still infinity. \
Adding 1 to the fourth element returns `NaN` because `NaN` is not a number and thus adding 1 is still not a number.
```{r}
x + 1
```

## 3.6.5 Q4
`c(NA, TRUE, NA)` \
The result of the operation on the first element is `NA` because it is not available and comparing two not available objects is also not available. \
The result of the operation on the second element is nothing because comparing two objects of length 0 returns a logical vector of length 0. \
The result of the operation on the third element is`TRUE` because infinity is equal to infinity in R. \
The result of the operation on the fourth element is `NA` because comparing two not a number values is not available.
```{r}
x == x
```