---
title: "HW5"
author: "Andrew Shao"
date: "2024-10-09"
output: pdf_document
header-includes: 
  - \usepackage{tikz}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 5.10.2 Q1
```{r}
library(r02pro)
library(ggplot2)

ggplot(data = sahp) +
  geom_density(aes(x = liv_area,
                   color = kit_qual),
               linetype = 'dashed')
```
Living area is skewed right for all the kitchen qualities. As kitchen quality improves the center of the living area distribution shifts to the right.

## 5.10.2 Q2
```{r}
ggplot(data = sahp) +
  geom_density(aes(x = kit_qual))
```
This plot is not informative since kit_qual is a categorical variable it doesn't make sense to use a density plot.
```{r}
library(forcats)
ggplot(data = sahp) +
  geom_bar(aes(x = fct_relevel(kit_qual,
                               c('Fair',
                                 'Average',
                                 'Good',
                                 'Excellent')
                               ))) +
  xlab('kit_qual')
```

## 5.11.6 Q1
```{r}
ggplot(data = sahp) +
  geom_boxplot(aes(x = '', y = liv_area))
```
Line in the middle:
```{r}
median(sahp$liv_area)
```
Lower hinge:
```{r}
quantile(sahp$liv_area, 0.25)
```
Upper hinge:
```{r}
quantile(sahp$liv_area, 0.75)
```
Lower whisker:
```{r}
quantile(sahp$liv_area, 0.25) - 1.5 * IQR(sahp$liv_area)
```
Upper whisker:
```{r}
quantile(sahp$liv_area, 0.75) + 1.5 * IQR(sahp$liv_area)
```

## 5.11.6 Q2
```{r}
ggplot(data = sahp) +
  geom_boxplot(aes(x = fct_relevel(kit_qual,
                               c('Fair',
                                 'Average',
                                 'Good',
                                 'Excellent')
                               ),
                   y = liv_area))
```
As kitchen quality improves the distribution of living area shifts in the positive direction.
  
## 5.11.6 Q3
```{r}
ggplot(data = sahp,
       aes(x = fct_relevel(kit_qual,
                           c('Fair',
                             'Average',
                             'Good',
                             'Excellent')
                           ),
           y = liv_area)) +
  geom_boxplot() +
  geom_point(stat = 'summary',
             fun = 'min',
             color = 'red') +
  geom_point(stat = 'summary',
             fun = 'max',
             color = 'blue') +
  geom_point(stat = 'summary',
             fun = 'mean',
             color = 'green')
```

## 5.11.6 Q4
```{r}
ggplot(data = sahp) +
  geom_boxplot(aes(x = fct_reorder(kit_qual,
                                   lot_area,
                                   mean,
                                   .desc = F
                                   ),
                   y = liv_area))
```

## 5.11.6 Q5
```{r}
ggplot(data = remove_missing(sahp, vars = 'oa_qual')) +
  geom_boxplot(aes(x = fct_relevel(kit_qual,
                               c('Fair',
                                 'Average',
                                 'Good',
                                 'Excellent')
                               ),
                   y = liv_area,
                   colour = oa_qual > 5))
```

## 5.12.3 Q1
```{r}
my_boxplot <- ggplot(data = sahp) +
  geom_boxplot(aes(y = liv_area))

my_boxplot
```

## 5.12.3 Q2
```{r}
my_boxplot + facet_wrap('bedroom')
```
As bedroom count increases so does the distribution of living area... which makes sense.

## 5.12.3 Q3
```{r}
my_boxplot + facet_grid(rows = vars(kit_qual),
                        cols = vars(central_air))
```
There is an empty plot for `kit_qual` value of `"Excellent"` and `central_air` value of `"N"`, this is because there is no observations within the dataset with that combination of values.