---
title: "In-class Assignment 6"
author: "Andrew Shao"
date: "2024-10-09"
output: pdf_document
header-includes: 
  - \usepackage{tikz}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 4.2.5 Q1
```{r}
my_data <- data.frame(word = c('one', 'two', NA, 'four', 'five'),
                      number = c(1, NA, 3, 4, 5),
                      letter = c('a', 'b', 'c', 'd', 'e'))
```
```{r}
setwd("C:/Users/as345/2182/2182")
if (!dir.exists('my_data')) {
  dir.create("my_data")
}

library(readr)
write_csv(my_data, 'my_data/my_data_no_name.csv', col_names = F)
```

## 4.2.5 Q2
```{r}
write_delim(my_data, 'my_data/my_data_na.csv', delim = '#', na = '999')
```

## 4.3.6 Q1
```{r}
my_data <- read_delim('my_data/my_data_na.csv', delim = '#', na = '999')
my_data
```

## 4.3.6 Q2
The column names should be `X1`, `X2`, and `X3`, because of the argument `col_names = FALSE` which tells the first row is not the column names and thus automatically assign default column names.
```{r}
d1 <- read_csv("x,y,z
          1,3,5
          2,4,6", col_names = FALSE)

d1
```

## 4.3.6 Q3
There will be only one column named `The first line` since R will assume the first line is the column names and since the default delimiter is `,` it will read the entire first line as just 1 string.
```{r}
d1 <- read_csv("The first line 
          The second line
          The third line
          x,y,z
          1,3,5")
d1
```