---
title: "In-class Assignment 9"
author: 'Andrew Shao (NetID: as13381)'
output: pdf_document
header-includes: 
  - \usepackage{tikz}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

### For Questions 1-5, we consider the relational data from the R package `nycflights13`.
```{r, include=FALSE}
library(tidyverse)
library(nycflights13)
```

### Question 1 (1 pt): Add the locations (i.e. the `lat` and `lon`) of the origin and destination to `flights`.
#### Answer: 
<!--Provide your answer and code below-->
```{r}
flights %>% 
  left_join(select(airports, faa, lat, lon), by = c('origin' = 'faa')) %>%
  rename_with(~paste0('origin_', .x), all_of(c('lat', 'lon'))) %>%
  left_join(select(airports, faa, lat, lon), by = c('dest' = 'faa')) %>%
  rename_with(~paste0('dest_', .x), all_of(c('lat', 'lon')))
```



### Question 2 (1 pt): Filter `flights` to only show flights with planes that have flown at least 100 flights.
#### Answer: 
<!--Provide your answer and code below-->
```{r}
flights %>% semi_join(flights %>% 
                        group_by(tailnum) %>% 
                        summarise(nflights = n()) %>% 
                        filter(nflights >= 100),
                      by = 'tailnum')
```



### Question 3 (1 pt): What does `anti_join(flights, airports, by = c("dest" = "faa"))` tell you? What does `anti_join(airports, flights, by = c("faa" = "dest"))` tell you?
#### Answer: 
<!--Provide your answer and code below-->
`anti_join(flights, airports, by = c("dest" = "faa"))` shows the flights whose destination is at an airport that is not included in the `airports` table. \
`anti_join(airports, flights, by = c("faa" = "dest"))` shows the airports in `airports` which there are no flights in the `flights` table that have them as destinations.
```{r}
anti_join(flights, airports, by = c("dest" = "faa"))
anti_join(airports, flights, by = c("faa" = "dest"))
```



### Question 4 (1 pt): Draw a plot for the average of departure-delay hours (`flights$dep_delay`) vs. the precipitation amount (`weather$precip`). 
#### Answer: 
<!--Provide your answer and code below-->
```{r}
flights %>%
  group_by(year, month, day, hour, origin) %>%
  summarise(avg_delay = mean(dep_delay, na.rm = T)) %>%
  inner_join(weather, by = c('year', 'month', 'day', 'hour', 'origin')) %>%
  ggplot() +
  geom_point(aes(x = avg_delay, y = precip))
```



### Question 5 (1 pt): Create data frames `flight1` and `flight2` using the following code. How many unique observations do these two data frames contain in total?
```{r Q5}
flights_daily <- flights%>%select(year:day, origin, dest)
flight1 <- flights_daily[1:1e5,]
flight2 <- flights_daily[5e4:2e5,]
```
#### Answer: 
<!--Provide your answer and code below-->
```{r}
nrow(union(flight1, flight2))
```