---
title: "In-class Assignment 11"
author: 'Andrew Shao (NetID: as13381)'
output: pdf_document
header-includes: 
  - \usepackage{tikz}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r, include=FALSE}
library(tidyverse)
library(stringr)
```

### Question 1 (1 pt): Use `str_length()` and `str_sub()` to extract the middle character from each string in the vector `x <- c("a", "abc", "abcd", "abcde", "abcdef")`. What will you do if a string has an even number of characters? (Hint: Base R function `floor()` or `ceiling()` may be useful here.)
**Answer:** 
<!--Provide your answer and code below-->
```{r}
x <- c("a", "abc", "abcd", "abcde", "abcdef")

str_sub(x, ceiling(str_length(x) / 2), ceiling(str_length(x) / 2))
```

### Question 2 (1 pt): `words` is a vector with each element being a word, which is available after you run `library(tidyverse)`. From the vector `words`, find the words that have the highest number of vowels?
**Answer:** 
<!--Provide your answer and code below-->
```{r}
words[sapply(str_extract_all(words, '[aeiou]'), length) == max(sapply(str_extract_all(words, '[aeiou]'), length))]
```
### Question 3 (1 pt): `sentences` is a vector with each element being a sentence, which is available after you run `library(tidyverse)`. Extract the first word from each sentence in `sentences`. Use `%>% head(100)` to output the result. 
**Answer:** 
<!--Provide your answer and code below-->
```{r}
str_split(sentences, ' ', simplify = T)[,1] %>% head(100)
```
### Question 4 (1 pt): In the vector `sentences`, find all words that contain an apostrophe. In each found word, separate out the pieces before and after the apostrophe. (Hint: See 14.4.3 Grouped Matches in the textbook.)
**Answer:** 
<!--Provide your answer and code below-->
```{r}
str_split(sentences, ' ', simplify = T)[str_split(sentences, ' ', simplify = T) %>% str_detect('\'')] %>%
  str_split('\'', simplify = T)
```
### Question 5 (1 pt): What are the five most common words in the vector `sentences`? (Hint: You may use `boundary("word")` with an appropriate function. Use the function `unlist()` to convert a list to a vector. The function `count()` may be useful here.)
**Answer:** 
<!--Provide your answer and code below-->
```{r}
sort(table(str_to_lower(unlist(str_split(sentences, boundary('word'))))), decreasing = T)[1:5]
```