---
title: "In-class Assignment 2"
author: "Andrew Shao"
date: "2024-09-19"
output: pdf_document
header-includes: 
  - \usepackage{tikz}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.4.4 Q1
`mix_1` will be `"integer"` because it is the sum of two `"integer"`s (int_1 and int_2 have "L" at the end) so using the coercion rule the storage type will be of the most complex type which is `"integer"`.\
`mix_2` will be `"double"` because it is the sum of an `"integer"` and a `"double"` (num_1 doesn't have the "L"), so the most complex type is `"double"` and we invoke the coercion rule like before.\
`mix_3` will be `"double"` because it is the quotient of two `"integer"`s that isn't a whole number so it must be stored as a `"double"`.\
`mix_4` will be `"character"` because it contains a `"double"` and a `"character"`, and `"character"` is the more complex type and we can invoke the coercion rule.\
`mix_5` will be `"double"` because it contains a `"double"` and a `"logical"`, and `"double"` is the more complex and we invoke the coercion rule.\
`mix_6` will be `"character"` because it contains a `"double"`, `"character"`, and `"logical"`, and `"character"` is the most complex of the three and we invoke the coercion rule.\

```{r}
int_1 <- 5L
int_2 <- 6L
num_1 <- 2
char_1 <- "pig"
logi_1 <- TRUE
mix_1 <- int_1 + int_2 
mix_2 <- int_1 + num_1
mix_3 <- int_1/int_2
mix_4 <- c(num_1, char_1)
mix_5 <- c(num_1, logi_1)
mix_6 <- c(num_1, char_1, logi_1)
```
```{r}
typeof(mix_1)
```
```{r}
typeof(mix_2)
```
```{r}
typeof(mix_3)
```
```{r}
typeof(mix_4)
```
```{r}
typeof(mix_5)
```
```{r}
typeof(mix_6)
```

## 2.4.4 Q2
`3 * logi_2 + logi_3` will result in a vector of storage type `"double"` with values 4, 1, and 4.\
`logi_2 - logi_3` will result in a vector of storage type `"double"` with values 0, -1, and 0.\
This is because the logical vectors will be treated as vectors of 0's and 1's in math operations.

## 2.5.3 Q1
```{r}
v1 <- c(7, 2, 4, 9, 7)
v2 <- c(6, 2, 8, 7, 9)
v3 <- 1:50
c(1, 2, 3, 4, 5)[v1 < v2]
```

## 2.5.3 Q2
```{r}
v2[v1 > 5]
```

## 2.5.3 Q3
```{r}
v3[v3 %% 7 == 0]
```

## 2.5.3 Q4
```{r}
v3[v3 %% 8 == 0] <- 100
v3
```

## 2.7.4 Q1
```{r}
exe <- c(2, 0, -3, 0, 5, 6)
sort(exe, decreasing = T)
```

## 2.7.4 Q2
```{r}
rank(exe, ties.method = 'min')[exe %in% c(2, 0)]
```

## 2.7.4 Q3
```{r}
order(exe)
```

## 2.13.6 Q1
```{r}
v1 <- seq(from = 1, to = 100, by = 3)
v2 <- sqrt(v1)
v1s <- v1[v1 > 30 & v1 < 60]
v1s
```

## 2.13.6 Q2
```{r}
v2[v1 < 20 | v1 > 50]
```

## 2.13.6 Q3
```{r}
a <- c(T, F, T, F)
b <- c(T, T, F, F)
xor(a, b) == (!a & b) | (a & !b)
```

## 2.15.3 Q1
```{r}
x <- rep(c(1, 2, NA), 3:5)
```
### a.
```{r}
summary(x)
```
```{r}
min(x, na.rm = T)
```
```{r}
quantile(x, 0.25, na.rm = T)
```
```{r}
median(x, na.rm = T)
```
```{r}
mean(x, na.rm = T)
```
```{r}
quantile(x, 0.75, na.rm = T)
```
```{r}
max(x, na.rm = T)
```
```{r}
sum(is.na(x))
```

### b.
```{r}
which(is.na(x))
```

### c.
```{r}
x_no_na <- na.omit(x)
x_no_na
```

### d.
```{r}
x[is.na(x)] <- median(x, na.rm = T)
x
```

## 2.15.3 Q2 
```{r}
y <- rep(c("N", 2, "A"), 5:3)
y[y %in% c('N', 'A')] <- NA
y
```
```{r}
which(is.na(y))
```